### 2019-09-22 papers/journal/2019_ACTA_Rebuttal_Rd2.txt

  Author response: We apologise for the confusing structure 
  of our text. To clarify this point we have added a new paragraph:

  "It is perhaps counter-intuitive that, beyond a certain degree 
  of edge-connectedness, any VTS graph can be implemented as a 
  biological system. This is because we have not placed any 
  constraints on the number of distinct SNAREs and other trafficking 
  molecules a cell contains. In practice there is an upper bound on the 
  number of molecules a cell can genetically encode. This question 
  ("What is the maximum edge-connectedness possible given a 
  certain number of SNAREs?") is complementary to the question we 
  have asked here ("What is the minimum edge-connectedness possible 
  given any number of SNARES operating under certain physical constraints?")."

  >> I also found the formulation of the main problems with 
  respect to connectedness somewhat confusing. For example, 
  in Sec. 5.1.1 and Sec. 5.1.2 the problems are formulated   
  as a search for VTS G such that G is not k-connected. 
  Yet, Sec. 5.1 specifies that "In case the formula is 
  satisfiable the procedure terminates and we report the 
  value of k", which seems contradictory.

  Authors response: In Sec. 5 we try to give a high-level 
  specification of the procedure, followed by a formal search 
  statement from a perspective of an efficient implementation. 
  This difference was only hinted but not explicitly stated. 
  We have now rewritten the parts of the text to address the 
  distinction between high-level overview and the implementation.

  "The implementation of the k-connectedness property is 
  challenging as it requires reasoning about all possible k − 1 
  edge removals. So instead of checking for the graph to be 
  k-connected, the encoding (and the implementation) checks 
  whether the graph is not k + 1-connected..."

  The use of "k" in the 6.1.6 k-connectivity constraints, 
  is overloaded. This subsection specifies how we encode 
  a graph that is not k-connected.

  We have also removed the specific search problems (as suggested 
  by Reviewer #3 in the first revision) for further apparent 
  accessibility of the subject matter.

### 2019-09-23 papers/journal/2019_ACTA_Rebuttal_Rd2.txt

  >> I also found the formulation of the main problems with 
  respect to connectedness somewhat confusing. For example, 
  in Sec. 5.1.1 and Sec. 5.1.2 the problems are formulated   
  as a search for VTS G such that G is not k-connected. 
  Yet, Sec. 5.1 specifies that "In case the formula is 
  satisfiable the procedure terminates and we report the 
  value of k", which seems contradictory.

  Authors response: In Sec. 5 we try to give a high-level 
  specification of the procedure, followed by a formal search 
  statement from a perspective of an efficient implementation. 
  This difference was only hinted but not explicitly stated. 
  We have now rewritten the parts of the text to address the 
  distinction between high-level overview and the implementation.

  "The implementation of the k-connectedness property is 
  challenging as it requires reasoning about all possible k − 1 
  edge removals. So instead of checking for the graph to be 
  k-connected, the encoding (and the implementation) checks 
  whether the graph is not k + 1-connected..."

  The use of "k" in the 6.1.6 k-connectivity constraints, 
  is overloaded. This subsection simply specifies how we encode 
  a graph that is not k-connected.

@@

  We have also removed the specific search problems (as suggested 
  by Reviewer #3 in the first revision) for further apparent 
  accessibility of the subject matter.

  >> Another point that should be revised further is around 
    clarifying the contributions of this paper. The response 
    document summarizes nicely that, together with additional 
    QBF encoding, "this paper brings together all the results 
    with extended experiments and more detailed expositions for 
    the benefit of both biological and computation community." 
    However, I did not find a similar concise statement of the 
    contributions in the text. In addition, some of the stated 
    advances over previous approaches seem incremental and minor. 
    For example, for the previous CBMC-based method it is stated 
    that "...the scalability of the method was insufficient. 
    The tool was able to handle the system with just eight 
    compartments". Yet, we learn that "A typical eukaryotic 
    cell contains 9 organelles" and even one of the methods 
    presented in this paper fails for some of the experimental 
    evaluations with 7 or more compartments. The contributions 
    of this work, perhaps in the integration and evaluation of
    several methods reported previously, should be stated clearly 
    and accurately in the main text.

  Authors response: We have stated that the key purpose of this 
  paper is one unified exposition of our work.

  "This paper brings together the above encodings with extended
   experiments and more detailed expositions for the benefit of
   both the biological and the computation communities..."

  We clarified the concerns of the reviewers and clearly 
  stated in the Sec 1. Introduction that the CBMC 
  method does not scale and the direct encoding does.

@@

  >> A final minor point is about the use of "Synthesis" as 
    an encoding in Fig 4 and the relevant explanations. 
    Since the other problems list verification techniques 
    and solvers (e.g. BMC, SMT, QBF) and no novel synthesis 
    algorithms were developed for this problem, it would be 
    more appropriate to specify QBF as the encoding used for 
    the "VTS synthesis" problem.

  Authors response: This is indeed the case. We have now 
  incorporated this information in Fig 4.

### 2019-09-24 papers/journal/conclusion.tex
Somya's Typo fixes.

\noindent 
We demonstrate in our work how local regulatory interactions between molecules of the VTS, translate into a global, observable properties of the VTS. 
%
Interactions between SNARE proteins and their regulators, which in its simplest form can be encoded in a pairing matrix, constrain the topological properties of the transport graph; all VTSs based on SNARE protein interactions must be 2-strongly connected. 
%
Additionally, we demonstrate how we can aid in predicting new vesicles given an incomplete VTS as input. 
%
Our approach demonstrates the power of SMT solvers in a novel application and the value of careful encoding of the problems into SMT queries. 
%
We manage to scale the tool up to the size that makes the tool biologically relevant.
%
%To conclude, we have established a connection between vesicle transport graph connectedness and underlying rules of SNARE pairing and regulation. 
%%
%The constraints implied by SNARE protein properties translates into
%mathematically elegant properties of the transport graph, namely its connectivity.
%%
%Our approach demonstrates the power of SMT solvers in a {\em novel} application and the value of careful encoding of the problems into SMT queries. 
%We manage to scale the tool up to the size that makes the tool biologically relevant.
%

However, the problem of searching for all $k$-connected VTSs induces quantifier alternation in the encodings. 
%
Therefore, a simple call to SAT solver may not work efficiently. 
%
We used QBF solvers for such queries. Our connectivity results contain many predictions which can in principle be tested through cell-biological experiments. 
%
For any input VTS, our tool outputs multiple possible solutions in the form of predictions of undiscovered vesicles and cargo molecules on specific vesicles. 
%
Such predictions could be used to guide biological experiments. Since a VTS that is close to completion is expected to have fewer alternative solutions to test, we expect that testing the predictions of our tool would become more tractable as more and more vesicles are discovered. 
%
One may also be interested in counting the number of graphs that satisfy a given property. 
%
Exact counting of the graphs using SAT solvers may prove to be very difficult. We are planning to employ approximate counting methods for finding the solution.

We also presented encodings of the synthesis problems that may arise from VTSs.
%
We demonstrated that our tool based on the encodings
scale up to the relevant sizes of the VTSs for some synthesis queries.
%
Our tool timed out on larger examples.
%
We are working to improve the performance of our tool.
%
%
With the tool, we made an attempt to present multiple new testable predictions. 
%
We will take this tool to the biologists and develop wet experiments that may
validate some synthesis results from the tool.
%
%These predictions may very likely increase to hasten discoveries in this field alongside pushing forward the use of formal methods in the biological field.
%

Since our current knowledge of VTSs is not extensive and there are many missing parts,
there is a huge space of potential predictions.
%
We also developing methods to prioritize predictions considering addition biological
information and using a refined model.
%
For example, in this work, we have only analyzed the VTS as static graphs. A major aspect of the biological cell’s VTS that the current model is missing is \textit{compartment maturation}, imbalances in incoming and outgoing molecular flux on vesicles.
%
In the future, we will study the dynamic behaviors of VTSs.

>>>>

\noindent 
We demonstrate in our work how local regulatory interactions between molecules of the VTS, translate into global, observable properties of the VTS. 
%
Interactions between SNARE proteins and their regulators, which in its simplest form can be encoded in a pairing matrix, constrain the topological properties of the transport graph; all VTSs based on SNARE protein interactions must be 2-strongly connected. 
%
Additionally, we demonstrate how we can aid in predicting new vesicles given an incomplete VTS as input. 
%
Our approach demonstrates the power of SMT solvers in a novel application and the value of careful encoding of the problems into SMT queries. 
%
We manage to scale the tool up to the size that makes it biologically relevant.
%
%To conclude, we have established a connection between vesicle transport graph connectedness and underlying rules of SNARE pairing and regulation. 
%%
%The constraints implied by SNARE protein properties translates into
%mathematically elegant properties of the transport graph, namely its connectivity.
%%
%Our approach demonstrates the power of SMT solvers in a {\em novel} application and the value of careful encoding of the problems into SMT queries. 
%We manage to scale the tool up to the size that makes the tool biologically relevant.
%

However, the problem of searching for all $k$-connected VTSs induces quantifier alternation in the encodings. 
%
Therefore, a simple call to SAT solvers may not work efficiently. 
%
We used QBF solvers for such queries. Our connectivity results contain many predictions which can in principle be tested through cell-biological experiments. 
%
For any input VTS, our tool outputs multiple possible solutions in the form of predictions of undiscovered vesicles and cargo molecules on specific vesicles. 
%
Such predictions could be used to guide biological experiments. Since a VTS that is close to completion is expected to have fewer alternative solutions to test, we expect that testing the predictions of our tool would become more tractable as more and more vesicles are discovered. 
%
One may also be interested in counting the number of graphs that satisfy a given property. 
%
Exact counting of the graphs using SAT solvers may prove to be very difficult. We are planning to employ approximate counting methods for finding the solution.

We also presented encodings of the synthesis problems that may arise from VTSs.
%
We demonstrated that our tool based on encodings
scales up to the relevant sizes of the VTSs for some synthesis queries.
%
Our tool timed out on larger examples, and we are working to improve its performance.
%
%
With the tool, we have made an attempt to present multiple new testable predictions. 
%
We plan to take this tool to biologists and develop wet experiments that may
validate some of our synthesis results.
%
%These predictions may very likely increase to hasten discoveries in this field alongside pushing forward the use of formal methods in the biological field.
%

Since our current knowledge of VTSs is not extensive and there are many missing parts,
there is a huge space of potential predictions.
%
We are also developing methods to prioritize predictions considering additional biological
information and using a refined model.
%
For example, in this work, we have only analyzed VTSs as static graphs. A major aspect of the biological cell’s VTS that the current model is missing is \textit{compartment maturation} that arises from imbalances in incoming and outgoing molecular flux on vesicles.
%
In the future, we will study dynamic behaviors of VTSs.

### 2019-09-24 papers/journal/problem.tex
Updated draft
Somya's Typo fixes.

\noindent %\improve{The weakest part of the paper; Need a serious rewrite!}
We analyze the model against our desired properties.
%
%Since in biology there are no declared specifications or properties, we conjecture some properties that appear to be 
%true in  relevant inall the known biological systems.
%
In this paper, we choose graph connectedness to be a key property.
%
We categorize the search problem into three different categories (Fig.~\ref{fig:vts-search}), search problem for the existential condition, search problem for the universal condition, and performing synthesis on the incomplete VTS. 
%
We will present two encodings for existential condition and one each for universal condition and synthesis.

@@

\vspace{0.2cm}
\noindent We will also seek to answer two kinds of questions given the property.
% %
% In this paper,
% our conjecture relates graph-connectedness to a particular variation of SNARE pairing and
% regulation rules of the VTS.
% %
% We are interested in the following two type of properties of the transport graph:
\begin{enumerate}
\item {\em Existential condition}:
if there is a VTS that satisfies the connectedness property.

\item {\em Universal condition}:
  if there is a connectedness property that is satisfied by each valid VTS.
\end{enumerate}
The goal is to find the minimum graph connectivity properties for each of the condition.
% In our program, the rules VTS needs to follow 
% %
% are represented as constrained over the transport graphs (G), label on the nodes and edges and activity of those labels (determined by a Boolean function f). 
% %
% Each of these constrained functions is defined to be TRUE for a G and f: if and only if the corresponding condition holds for the given G and f. 
% %
% We use these functions to define the assertion that characterizes the corresponding condition to be checked. 
% %
% We will show in later sections how we encode these Boolean functions. 
\subsection{Search for existential condition} 
%
% To ensure that k-connectedness is an existential condition,
%
We need to find a $k$-connected graph that satisfies all the rules (and constraints) of the system. 
%
We also aim to find the minimum $k$. 
%
%Let's fix our conjecture: ``k-connectedness is a necessary condition for the system regulated by a Boolean function on the node and SNARE-SNARE inhibition on the edge". 
%
To find a $k$-connected graph we specify our property as a query for the existence
of a model for a conjunction of the formula representing stability, fusion and specific
graph connectivity constraint. 
%
% \begin{align*}
%   &\texttt{LabeledVTS} = \texttt{Stability} \, \land \, \texttt{Fusion} \land \, \texttt{MinConnectivity(k)}
%  \tag{E}\label{eq:existcond}
% \end{align*}
%
We will start with the value of $k = 1$ and check for the satisfiability of the formula.
%
In case the formula is satisfiable the procedure terminates and we report the value of $k$. 
%
If the formula is not satisfiable we increment the $k$ by $1$ and repeat the same procedure. 
%
In this way, we ensure that the reported $k$ is the minimum connectivity of the VTS that
satisfies the existential condition.
%

@@

We employ two techniques to examine the existential condition. 
%
First, using the BMC and second using SMT solving.
%
Throughout the paper, we will address these two subcategories as $\fbmc$ and $\fsmt$ search problems.
%
%Here, we present both the $\fbmc$ and $\fsmt$ search problem for the existential condition.
%
%\subsubsection{$\fbmc$ search problem.}
%For a given $k$, size $\nu$, molecule number $\mu$, and unwinding depth $w$,
%we are searching for well-structured, stable, and well-fused VTS
%$G = (\nodes,\mols,\edges,\nlabel,\pairs,\edgef,\nodef)$ such that
%$|\nodes| = \nu$, $|\mols| = \mu$, and $G$ is not $k$-connected, within $w$ depth of unwinding.    
%
%\subsubsection{$\fsmt$ search problem.}
%For a given $k$, size $\nu$, and molecule number $\mu$,
%we are searching for well-structured, stable, and well-fused VTS
%$G = (\nodes,\mols,\edges,\nlabel,\pairs,\edgef,\nodef)$ such that
%$|\nodes| = \nu$, $|\mols| = \mu$, and
%$G$ is not $k$-connected.

\subsection{Search for universal condition}
%
% Similarly, for the conjecture that k-connectedness is a universal condition for the VTS,
For some $k>0$,
we have to ensure that every graph of $k$-connected structure satisfies all
the rules of the system.
%
We also aim to find the minimum $k$. 
% We also have to ensure that the particular k is the least one. 
%
This condition can be specified as: for every $k$-connected graph, there exists a satisfiable assignment following the rules of the system. 
%
% \begin{align*}
%   & \forall G \, \texttt{Connectivity(k)} \implies  \exists
%                         f,p: \texttt{Stability} \, \land \, \texttt{Fusion}  
%   \tag{U}\label{eq:univcond}
% \end{align*}
%
We will use the same procedure as the existential condition to ensure that $k$ is minimized.
%

@@

For the efficient implementation of the specification instead of checking the satisfiability of the formula, we check whether the negation of the formula is unsatisfiable. 
%
That is to conclude that a given $k$-connectedness is not a universal condition we search for a $k$-connected graph that has no satisfying assignment to all the rules of the system.
%
%In case no such graph exists for a specific $k$ we report the $k$.
%%\ankit{This is a dummy template. need a rewrite.} 
%\subsubsection{$\fqbf$ search problem.} For a given $k$, size $\nu$, and molecule number $\mu$, we are searching for a $k$-connected graph such that there is no assignment that makes it a well-structured, stable, and well-fused VTS $G = (\nodes,\mols,\edges,\nlabel,\pairs,\edgef,\nodef)$ with $|\nodes| = \nu$, $|\mols| = \mu$.

@@

%\subsection{Problem Statements}
\subsection{Synthesis of VTS}
%
\noindent We also consider another variant of analysis of VTSs.
%
We will assume that we are given a VTS, whose all components
are not specified.
%
Our objective is to find the missing components.
%
The missing component can
% be in any of the components of VTS. 
%
%For example, 
be some undiscovered edges or nodes, or insufficient
knowledge about the presence of molecules in some part of the VTS.
%
To cover most of the likely variations of this missing information,
we have encoded the following variants of the VTS synthesis problem.

>>>>

\noindent %\improve{The weakest part of the paper; Need a serious rewrite!}
We analyze the model against our desired properties.
%
%Since in biology there are no declared specifications or properties, we conjecture some properties that appear to be 
%true in  relevant inall the known biological systems.
%
In this paper, we choose graph connectedness to be a key property.
%
%We categorize the search problem into three different categories (Fig.~\ref{fig:vts-search}), search problem for the existential condition, search problem for the universal condition, and performing synthesis on the incomplete VTS. 
%
%We will present two encodings for the existential condition and one each for the universal condition and synthesis.

@@

\vspace{0.2cm}
\noindent We seek to answer three kinds of questions given the property (Fig.~\ref{fig:vts-search}).
% %
% In this paper,
% our conjecture relates graph-connectedness to a particular variation of SNARE pairing and
% regulation rules of the VTS.
% %
% We are interested in the following two type of properties of the transport graph:
\begin{enumerate}
\item {\em Existential condition}:
if there is a VTS that satisfies the connectedness property.

\item {\em Universal condition}:
  if there is a connectedness property that is satisfied by each valid VTS.
\item{\em synthesis on incomplete VTS}:
    Given an incomplete VTS, if we can predict missing parts.
\end{enumerate}
The goal is to find the minimum graph connectivity properties for each of the condition. We will present two encodings for the existential condition and one each for the universal condition and synthesis.
% In our program, the rules VTS needs to follow 
% %
% are represented as constrained over the transport graphs (G), label on the nodes and edges and activity of those labels (determined by a Boolean function f). 
% %
% Each of these constrained functions is defined to be TRUE for a G and f: if and only if the corresponding condition holds for the given G and f. 
% %
% We use these functions to define the assertion that characterizes the corresponding condition to be checked. 
% %
% We will show in later sections how we encode these Boolean functions. 
\subsection{Search for existential condition} 
%
% To ensure that k-connectedness is an existential condition,
%
We need to find a $k$-connected graph that satisfies all the rules (and constraints) of the system. 
%
We also aim to find the minimum $k$. 
%
%Let's fix our conjecture: ``k-connectedness is a necessary condition for the system regulated by a Boolean function on the node and SNARE-SNARE inhibition on the edge". 
%
To find a $k$-connected graph we specify our property as a query for the existence
of a model for a conjunction of the formula representing stability, fusion and the specific
graph connectivity constraint. 
%
% \begin{align*}
%   &\texttt{LabeledVTS} = \texttt{Stability} \, \land \, \texttt{Fusion} \land \, \texttt{MinConnectivity(k)}
%  \tag{E}\label{eq:existcond}
% \end{align*}
%
We will start with the value of $k = 1$ and check for the satisfiability of the formula.
%
In case the formula is satisfiable the procedure terminates and we report the value of $k$. 
%
If the formula is not satisfiable we increment $k$ by $1$ and repeat the same procedure. 
%
In this way, we ensure that the reported $k$ is the minimum connectivity of the VTS that
satisfies the existential condition.
%

@@

We employ two techniques to examine the existential condition: first, using BMC and second using SMT solving.
%
Throughout the paper, we will address these two subcategories as $\fbmc$ and $\fsmt$ search problems.
%
%Here, we present both the $\fbmc$ and $\fsmt$ search problem for the existential condition.
%
%\subsubsection{$\fbmc$ search problem.}
%For a given $k$, size $\nu$, molecule number $\mu$, and unwinding depth $w$,
%we are searching for well-structured, stable, and well-fused VTS
%$G = (\nodes,\mols,\edges,\nlabel,\pairs,\edgef,\nodef)$ such that
%$|\nodes| = \nu$, $|\mols| = \mu$, and $G$ is not $k$-connected, within $w$ depth of unwinding.    
%
%\subsubsection{$\fsmt$ search problem.}
%For a given $k$, size $\nu$, and molecule number $\mu$,
%we are searching for well-structured, stable, and well-fused VTS
%$G = (\nodes,\mols,\edges,\nlabel,\pairs,\edgef,\nodef)$ such that
%$|\nodes| = \nu$, $|\mols| = \mu$, and
%$G$ is not $k$-connected.

\subsection{Search for universal condition}
%
% Similarly, for the conjecture that k-connectedness is a universal condition for the VTS,
For some $k>0$,
we have to ensure that every $k$-connected graph satisfies all
the rules of the system.
%
We also aim to find the minimum such $k$. 
% We also have to ensure that the particular k is the least one. 
%
This condition can be specified as: for every $k$-connected graph, there exists a satisfiable assignment following the rules of the system. 
%
% \begin{align*}
%   & \forall G \, \texttt{Connectivity(k)} \implies  \exists
%                         f,p: \texttt{Stability} \, \land \, \texttt{Fusion}  
%   \tag{U}\label{eq:univcond}
% \end{align*}
%
We will use the same procedure as the existential condition to ensure that $k$ is minimized.
%

@@

For an efficient implementation of the specification, instead of checking the satisfiability of the formula, we check whether the negation of the formula is unsatisfiable. 
%
That is, to conclude that a given $k$-connectedness is not a universal condition, we search for a $k$-connected graph that has no satisfying assignment to all the rules of the system.
%
%In case no such graph exists for a specific $k$ we report the $k$.
%%\ankit{This is a dummy template. need a rewrite.} 
%\subsubsection{$\fqbf$ search problem.} For a given $k$, size $\nu$, and molecule number $\mu$, we are searching for a $k$-connected graph such that there is no assignment that makes it a well-structured, stable, and well-fused VTS $G = (\nodes,\mols,\edges,\nlabel,\pairs,\edgef,\nodef)$ with $|\nodes| = \nu$, $|\mols| = \mu$.

@@

%\subsection{Problem Statements}
\subsection{Synthesis of VTS}
%
\noindent We consider another variant of analysis of VTSs.
%
We will assume that we are given a VTS, all of whose components
are not specified.
%
Our objective is to find the missing components.
%
The missing component can
% be in any of the components of VTS. 
%
%For example, 
be undiscovered edges or nodes, or insufficient
knowledge about the presence of molecules in some part of the VTS.
%
To cover most of the likely variations of this missing information,
we have encoded the following variants of the VTS synthesis problem.

### 2019-09-24 papers/journal/related.tex
Somya's Typo fixes.

\noindent The previous analysis of VTS used simulation and statistical reasoning~\cite{mani2016stacking}. 
%
Such analysis is far from complete and can only make statistical claims about VTS based on a subset of samples. 
%
The properties of VTS discussed in this paper require precise analysis for all possible rules and variations, that make formal methods a suitable alternative.

The SAT solvers are used to solve a wide range of problems including
%
software and hardware verification~\cite{biere1999symbolic1, biere1999symbolic2, bjesse2001finding, velev2003effective}, equivalence checking~\cite{goldberg2001using}, automatic test case generation~\cite{stephan1996combinational},
AI planning~\cite{kautz1996pushing} and scheduling~\cite{gomes1998randomization}. 
%
The enormous progress in the performance of SAT and SMT solvers
%Improvement in SAT and SMT solvers
in the past decades have led to their use in modeling and understanding complex biological systems~\cite{heule2010exact,yordanov2013smt,mangla2010timing}. 
%
Many of which involve reasoning about graphs and networks rules~\cite{guerra2012reasoning,chin2008biographe} with possible exhaustive search.
% and hence can be reduced to solving a SAT/SMT question. 
%

One of the prominent areas of application of SAT and SMT solvers is modeling and understanding the gene regulatory networks (GRN)~\cite{giacobbe2015model,rosenblueth2014inference, yordanov2016method, dunn2014defining, paoletti2014analyzing, koksal2013synthesis}.
%
The gene regulatory network question is encoded either directly as a Boolean satisfiability problem or in addition with the combinations of background theories.
%
The generated formula is then solved with the help of a SAT solver. 
%
The~\cite{rosenblueth2014inference} uses symbolic model checking approach on the qualitative
models of regulatory networks.
%
Whereas,~\cite{dunn2014defining}  uses symbolic, SAT-based approach on the Boolean network model. 
%
The~\cite{giacobbe2015model, yordanov2016method} presents an SMT based encoding of the problem based on the presented Boolean model.
%

We have extended the application of SAT and SMT solvers to a comparatively more complex biological transport network~\cite{mani2016stacking} namely VTS. 
%
We have presented both SAT and SMT based encoding for our Boolean model of the VTS.
%
We have also introduced the use of QBF solvers and it's encoding in this context.
% 
%To our best knowledge, this series of work is the first application
%of formal methods to the vesicle traffic networks.

In recent years, there has been a wide range of methods
developed for similar synthesis problems~\cite{sketch,sygus,exampleSynth}.
%
They range from filling gaps an implementation of C programs from the pool of template predicates to learn a program from the example
runs of the program.
%
In the course of developing such methods,
the background technology, i.e. solving of quantified
constraints has been evolving rapidly~\cite{lonsing2010depqbf,z3Quant}.
%

There has been some work in applying synthesis technique in biology. 
%
Recent work by \cite{fisher2015synthesising} synthesize executable gene regulatory networks from single-cell gene expression data.
%
In \cite{shavit2016automated} synthesis of switching GRN (with changing topology) is performed.
%
The~\cite{fisher2015synthesising} uses constraint (satisfiability) solving techniques whereas \cite{shavit2016automated} uses SMT to perform synthesis. 
%
Synthesis technique is also used in optimal synthesis for chemical reaction networks~\cite{cardelli2017syntax}.
%
%
The work uses a template-guided approach in addition to using SMT over ODE. In our case, queries contain quantifiers so we have employed QBF solving with Z3 for solving the synthesis problem. 
%
To our best knowledge, this is the first application of synthesis in VTS.
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

>>>>

\noindent Previous analysis of VTS used simulation and statistical reasoning~\cite{mani2016stacking}. 
%
Such analysis is far from complete and can only make statistical claims about the VTS based on a subset of samples. 
%
The properties of VTS discussed in this paper require precise analysis for all possible rules and variations, that make formal methods a suitable alternative.

SAT solvers are used to solve a wide range of problems including
%
software and hardware verification~\cite{biere1999symbolic1, biere1999symbolic2, bjesse2001finding, velev2003effective}, equivalence checking~\cite{goldberg2001using}, automatic test case generation~\cite{stephan1996combinational},
AI planning~\cite{kautz1996pushing} and scheduling~\cite{gomes1998randomization}. 
%
Enormous progress in the performance of SAT and SMT solvers
%Improvement in SAT and SMT solvers
in the past decades have led to their use in modeling and understanding complex biological systems~\cite{heule2010exact,yordanov2013smt,mangla2010timing}. 
%
Many of these applications involve reasoning about graphs and networks rules~\cite{guerra2012reasoning,chin2008biographe} with possible exhaustive search.
% and hence can be reduced to solving a SAT/SMT question. 
%

One of the prominent areas of application of SAT and SMT solvers is modeling and understanding gene regulatory networks (GRN)~\cite{giacobbe2015model,rosenblueth2014inference, yordanov2016method, dunn2014defining, paoletti2014analyzing, koksal2013synthesis}.
%
The gene regulatory network question is encoded either directly as a Boolean satisfiability problem or in addition with combinations of background theories.
%
The generated formula is then solved with the help of a SAT solver. 
%
~\cite{rosenblueth2014inference} uses a symbolic model checking approach on qualitative
models of regulatory networks.
%
Whereas,~\cite{dunn2014defining} uses a symbolic, SAT-based approach on the Boolean network model. 
%
~\cite{giacobbe2015model, yordanov2016method} presents an SMT based encoding of the problem based on the presented Boolean model.
%

We have extended the application of SAT and SMT solvers to a comparatively more complex biological transport network~\cite{mani2016stacking}, namely VTS. 
%
We have presented both SAT and SMT based encodings for our Boolean model of the VTS.
%
We have also introduced the use of QBF solvers and it's encoding in this context.
% 
%To our best knowledge, this series of work is the first application
%of formal methods to the vesicle traffic networks.

In recent years, a wide range of methods
have been developed for similar synthesis problems~\cite{sketch,sygus,exampleSynth}.
%
They range from filling gaps in implementation of C programs from the pool of template predicates, to learning a program from the example
runs of the program.
%
In the course of developing such methods,
the background technology, i.e. solving of quantified
constraints has been evolving rapidly~\cite{lonsing2010depqbf,z3Quant}.
%

There has been some work in applying synthesis technique in biology. 
%
Recent work by \cite{fisher2015synthesising} synthesize executable gene regulatory networks from single-cell gene expression data.
%
In \cite{shavit2016automated} synthesis of switching GRN (with changing topology) is performed.
%
~\cite{fisher2015synthesising} uses constraint (satisfiability) solving techniques, whereas \cite{shavit2016automated} uses SMT to perform synthesis. 
%
A synthesis technique is also used in optimal synthesis of chemical reaction networks~\cite{cardelli2017syntax}.
%
%
The work uses a template-guided approach in addition to using SMT over ODE. In our case, queries contain quantifiers so we have employed QBF solving with Z3 for solving the synthesis problem. 
%
To our best knowledge, this is the first application of synthesis in VTS.
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:
